Are environmental and geographic effective surrogates for genetic variation in conservation planning?

### 2016-01-20 article/article.Rmd
fixes to ms + removed unsampled cells from analysis

## Methods
### Study area
To address the aims of the study, we utilized species distribution and genomic data collected by the IntraBioDiv project [@r451; see @r452 for methodological details]. This dataset was well suited for this study, because it provides spatially explicit genomic data for a multitude of species across a broad geographic area and over large environmental gradients. As part of this project, data for 27 alpine plant species were collected across the European Alps according to a grid (22.3 km $\times$ 25 km; Figure S1). Briefly, in every second cell, plant samples were collected from three individuals for each species present inside the cell. Samples were genotyped using amplified fragment length polymorphisms [AFLP; @r453]. Matrices denoting the presence/absence of polymorphisms at loci were constructed independently for each species. Thus for each second grid cell, for each of the species in the grid cell, the dataset contains information describing the genomic variation of individuals inside the grid cell. We used these grid cells as planning units to develop prioritizations to investigate the effectiveness of surrogates for the (AFLP) genetic variation.

### Surrogate data
We explored the effectiveness of climatic and geographic surrogates. To describe variation in the location of each grid cell, we projected the grid into an equi-distant coordinate system ($\texttt{ESRI:102031}$) and used the centroid of each cell. To describe the climatic variation among each cell, we obtained twenty-one bioclimatic variables [approx. 1 km resolution; @r33]. These layers and the grid cells were projected into an equal-area coordinate system ($\texttt{ESRI:102014}$). The bioclimatic variables were clipped to the extent of the grid and subject to a principle components analysis (PCA; Table S1). The first two principle components (PCs) cumulatively explained `r round(pca.DF[[4]][2],1)` % of the variation and were used to generate two new layers. We calculated the average value for each PC layer in each grid cell, and used these values to characterize the main themes of climatic variation among the grid cells (Figure S2).

@@

First, we grouped individuals into genetic lineages for each species (Figure S2). The purpose of this step was to avoid confounding variation between populations attributed to genetic history instead of selection pressures. We used discriminant analyses of principle components [DAPC; @r467] to assign individuals to populations [implemented in $\texttt{adegenet}$ R package; @r468]. Briefly, for each species, we transformed the binary loci data using a principle components analysis (PCA) and obtained the principle components required to explain at least `r adegenet.params.LST[[MODE]]$perc.pca` % variation. We subject these components to multiple $k$-means clustering models with varying numbers of assumed populations (\textit{k} = 1--`r adegenet.params.LST[[MODE]]$max.n.clust`; `r adegenet.params.LST[[MODE]]$n.iter` iterations; `r adegenet.params.LST[[MODE]]$n.start` random starts). We then  applied Ward's clustering method to the Bayesian Information Criterion (BIC) statistics associated with each model to partition the models into two groups, and selected the model before the first group switch (as recomended in the \texttt{adegenet} user manual). The populations members associated with these models were used for subsequent analysis.

We used $\texttt{Bayescan}$ [version 2.1; @r456] to identify loci under selection for each species (Table S2). For each species, $\texttt{Bayescan}$ was run using the population memberships identified in the previous analysis (1:1 prior odds; `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r bayescan.params.LST[[MODE]]$burn` burn-in iterations; `r bayescan.params.LST[[MODE]]$n` post-burn-in iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations). Following guidelines in the $\texttt{Bayescan}$ manual, we omitted loci where the global frequency of the minor allele was $\geq 0.05$ to avoid false-positives. To ensure convergence, we ran `r bayescan.params.LST[[MODE]]$reps` replicates per species using a suitable false discovery rate [FDR; $\leq `r bayescan.params.LST[[MODE]]$fdr`$; @r466].

After classifying loci as adaptive or neutral, we mapped the main gradients of the adaptive (if present) and neutral genetic variation for each species (Figures S4--S29). For each species, for each type of genetic variation present, we used Gower distances [@r464] to express the genomic differences between individuals [using the $\texttt{cluster}$ R package to accommodate sparsity; @r457]. These distances were subject to multiple non-metric multi-dimensional scaling analyses using the minimum number of dimensions required to acheive an adequate stress value  [$\leq$ `r mds.params.LST[[MODE]]$max.stress`; `r mds.params.LST[[MODE]]$trymax` random starts; Table S3; implemented in the $\texttt{vegan}$ R package; @r458]. Since each grid cell had up to three samples per species, we used the average of the ordinated values associated with the samples in each grid cell to characterize the genomic variation among the grid cells.

>>>>

## Methods
### Study area
To address the aims of the study, we utilized species distribution and genomic data collected by the IntraBioDiv project [@r451; see @r452 for methodological details]. This dataset was well suited for this study, because it provides spatially explicit genomic data for a multitude of species across a broad geographic area and over large environmental gradients. As part of this project, data for 27 alpine plant species were collected across the European Alps according to a grid (22.3 km $\times$ 25 km; Figure S1). Briefly, in every second cell, plant samples were collected from three individuals for each species present inside the cell. Cells that did not contain any plant samples were omitted from analysis. Samples were genotyped using amplified fragment length polymorphisms [AFLP; @r453]. Matrices denoting the presence/absence of polymorphisms at loci were constructed independently for each species. Thus for each second grid cell, for each of the species in the grid cell, the dataset contains information describing the genomic variation of individuals inside the grid cell. We used these grid cells as planning units to develop prioritizations to investigate the effectiveness of surrogates for the (AFLP) genetic variation.

### Surrogate data
We explored the effectiveness of climatic and geographic surrogates. To describe variation in the location of each grid cell, we projected the grid into an equi-distant coordinate system ($\texttt{ESRI:102031}$) and used the centroid of each cell. To describe the climatic variation among each cell, we obtained 21 bioclimatic variables [approx. 1 km resolution; @r33]. These layers and the grid cells were projected into an equal-area coordinate system ($\texttt{ESRI:102014}$). The bioclimatic variables were clipped to the extent of the grid and subject to a principle components analysis (PCA; Table S1). The first two principle components (PCs) cumulatively explained `r round(pca.DF[[4]][2],1)` % of the variation and were used to generate two new layers. We calculated the average value for each PC layer in each grid cell, and used these values to characterize the main themes of climatic variation among the grid cells (Figure S2).

@@

First, we grouped individuals into genetic lineages for each species (Figure S2). The purpose of this step was to avoid misclassifying loci as adaptive. We used discriminant analyses of principle components [DAPC; @r467] to assign individuals to populations [implemented in $\texttt{adegenet}$ R package; @r468]. Briefly, for each species, we transformed the binary loci data using a principle components analysis (PCA) and obtained the principle components required to explain at least `r adegenet.params.LST[[MODE]]$perc.pca` % variation. We subject these components to multiple \textit{k}-means clustering models with varying numbers of assumed populations (\textit{k} = 1--`r adegenet.params.LST[[MODE]]$max.n.clust`; `r adegenet.params.LST[[MODE]]$n.iter` iterations; `r adegenet.params.LST[[MODE]]$n.start` random starts). We then  applied Ward's clustering method to the Bayesian Information Criterion (BIC) statistics associated with each model to partition the models into two groups, and selected the model before the first group switch (as recomended in the \texttt{adegenet} user manual). The populations members associated with these models were used for subsequent analysis.

We used $\texttt{Bayescan}$ [version 2.1; @r456] to identify loci under selection for each species (Table S2). For each species, $\texttt{Bayescan}$ was run using the population memberships identified in the previous analysis (1:1 prior odds; `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r bayescan.params.LST[[MODE]]$burn` burn-in iterations; `r bayescan.params.LST[[MODE]]$n` post-burn-in iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations). We omitted loci where the global frequency of the minor allele was $\geq 0.05$ to avoid false-positives (following guidelines in the $\texttt{Bayescan}$ manual). To ensure convergence, we ran `r bayescan.params.LST[[MODE]]$reps` replicates per species using a suitable false discovery rate [FDR; $\leq `r bayescan.params.LST[[MODE]]$fdr`$; @r466].

After classifying loci as adaptive or neutral, we mapped the main gradients of the adaptive (if detected) and neutral genetic variation for each species (Figures S4--S29). For each species, for each type of genetic variation detected, we used Gower distances [@r464] to express the genomic differences between individuals [using the $\texttt{cluster}$ R package to accommodate sparsity; @r457]. These distances were subject to multiple non-metric multi-dimensional scaling analyses using the minimum number of dimensions required to acheive an adequate stress value  [$\leq$ `r mds.params.LST[[MODE]]$max.stress`; `r mds.params.LST[[MODE]]$trymax` random starts; Table S3; implemented in the $\texttt{vegan}$ R package; @r458]. Since each grid cell had up to three samples per species, we used the average of the ordinated values associated with the samples in each grid cell to characterize the genomic variation among the grid cells.

### 2016-01-20 article/figures.Rmd
fixes to ms + removed unsampled cells from analysis

```{r, echo=FALSE, fig.width=8, fig.height=3, fig.cap='Multi-species prioritizations. Panel (a) shows the prioritization generated for using just amount-based targets. Panel (b) shows the prioritization generated using amount-based and surrogate based targets. Panel (c) shows the prioritization generated using amount-based and genetic-based targets.'}
# download basemap
data(countriesHigh)
countries.FPLY <- countriesHigh[
	countriesHigh$ADMIN %in% c(
		'Italy', 'Switzerland', 'France', 'Austria',
		'Germany', 'Slovenia', 'Croatia', 'Hungary',
		'Monaco', 'Germany'
	)
,] %>% spFortify
# prepare data for plotting
multi.spp.grid.FPLY <- grid.PLY
for (i in seq_along(multi.spp.prioritisations))
	multi.spp.grid.FPLY@data[[paste0('v',i)]] <- selections(multi.spp.prioritisations[[i]])
multi.spp.grid.FPLY <- spFortify(multi.spp.grid.FPLY)
# make maps
do.call(
	grid.arrange,
	append(
		llply(
			seq_along(multi.spp.prioritisations),
			function(i) {
				ggplot() +
					geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
						fill='grey20', color='grey80') +
					geom_polygon(data=multi.spp.grid.FPLY, aes_string(x='long', y='lat', 
						group='group', fill=paste0('v',i)),
						alpha=0.8, color='grey10') +
					guides(fill=guide_legend(title=' ')) +
					theme_classic() +
					theme(axis.ticks=element_blank(), axis.text=element_blank(),
						plot.margin=unit(c(0,0,0,0),'cm'), axis.line=element_blank(),
						legend.position='none') +
					coord_cartesian(
						xlim=buffered.range(multi.spp.grid.FPLY$long, 0.05),
						ylim=buffered.range(multi.spp.grid.FPLY$lat, 0.05)
					) +
					xlab('') +
					ylab('') +
					annotate('text',
						x=min(multi.spp.grid.FPLY$long)+diff(range(multi.spp.grid.FPLY$long)*0.1),
						y=min(multi.spp.grid.FPLY$lat)+diff(range(multi.spp.grid.FPLY$lat)*1.01),
						label=letters[i], hjust=1, vjust=1, color='white', size=8)
			}
		),
		list(nrow=1)
	)
)
```

>>>>

```{r, echo=FALSE, fig.width=8, fig.height=3, fig.cap='Multi-species prioritizations. Panel (a) shows the prioritization generated for using just amount-based targets. Panel (b) shows the prioritization generated using amount-based and surrogate based targets. Panel (c) shows the prioritization generated using amount-based and genetic-based targets.'}
# download basemap
data(countriesHigh)
countries.FPLY <- countriesHigh[
	countriesHigh$ADMIN %in% c(
		'Italy', 'Switzerland', 'France', 'Austria',
		'Germany', 'Slovenia', 'Croatia', 'Hungary',
		'Monaco', 'Germany', 'Slovakia', 'Czech Republic'
	)
,] %>% spFortify
# prepare data for plotting
multi.spp.grid.FPLY <- grid.PLY
for (i in seq_along(multi.spp.prioritisations))
	multi.spp.grid.FPLY@data[[paste0('v',i)]] <- selections(multi.spp.prioritisations[[i]])
multi.spp.grid.FPLY <- spFortify(multi.spp.grid.FPLY)
# make maps
do.call(
	grid.arrange,
	append(
		llply(
			seq_along(multi.spp.prioritisations),
			function(i) {
				ggplot() +
					geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
						fill='grey20', color='grey80') +
					geom_polygon(data=multi.spp.grid.FPLY, aes_string(x='long', y='lat', 
						group='group', fill=paste0('v',i)),
						alpha=0.8, color='grey10') +
					guides(fill=guide_legend(title=' ')) +
					theme_classic() +
					theme(axis.ticks=element_blank(), axis.text=element_blank(),
						plot.margin=unit(c(0,0,0,0),'cm'), axis.line=element_blank(),
						legend.position='none') +
					coord_cartesian(
						xlim=buffered.range(multi.spp.grid.FPLY$long, 0.05),
						ylim=buffered.range(multi.spp.grid.FPLY$lat, 0.05)
					) +
					xlab('') +
					ylab('') +
					annotate('text',
						x=min(multi.spp.grid.FPLY$long)+diff(range(multi.spp.grid.FPLY$long)*0.1),
						y=min(multi.spp.grid.FPLY$lat)+diff(range(multi.spp.grid.FPLY$lat)*1.01),
						label=letters[i], hjust=1, vjust=1, color='white', size=8)
			}
		),
		list(nrow=1)
	)
)
```

### 2016-01-20 article/supporting_information.Rmd
fixes to ms + removed unsampled cells from analysis

---
fontsize: 11pt
documentclass: article
bibliography: Endnote_lib.bib
csl: reference-style.csl
output:
  rmarkdown::pdf_document:
    fig_caption: yes
    includes:
      in_header: [preamble-latex.tex, preamble-latex2.tex]
---
## Supporting Information
All data and code is stored in an online repository (\url{www.github.com/paleo13/genetic.surrogates}) to replicate analyses in this study. See the \textit{README} file for instructions.

@@

# download basemap
data(countriesHigh)
countries.FPLY <- countriesHigh[
	countriesHigh$ADMIN %in% c(
		'Italy', 'Switzerland', 'France', 'Austria',
		'Germany', 'Slovenia', 'Croatia', 'Hungary',
		'Monaco', 'Germany'
	)
,] %>% spFortify
grid.FPLY <- spFortify(grid.PLY)
```

>>>>

---
fontsize: 11pt
documentclass: article
bibliography: Endnote_lib.bib
csl: reference-style.csl
output:
  rmarkdown::pdf_document:
    fig_caption: yes
    includes:
      in_header: [preamble-latex.tex, preamble-latex3.tex]
---
## Supporting Information
All data and code is stored in an online repository (\url{www.github.com/paleo13/genetic.surrogates}) to replicate analyses in this study. See the \textit{README} file for instructions.

@@

# download basemap
data(countriesHigh)
countries.FPLY <- countriesHigh[
	countriesHigh$ADMIN %in% c(
		'Italy', 'Switzerland', 'France', 'Austria',
		'Germany', 'Slovenia', 'Croatia', 'Hungary',
		'Monaco', 'Germany', 'Slovakia', 'Czech Republic'
	)
,] %>% spFortify
grid.FPLY <- spFortify(grid.PLY)
```
